Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-07-26 analysis/paper/paper.Rnw
replaced file.path calls with here::here

<<readRDS-2>>=
pleio_mixed_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_pleio.rds")))
pleio_local_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_pleio_local.rds")))
pleio_nonlocal_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
med_list <- readRDS(here("analysis", "data", paste0("Chr", hot_chr, "_mediations.rds")))
@

<<load_attie>>=
load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
@

<<make_local_annot-2>>=
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
@

<<make_nonlocal_annot-2>>=
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  inner_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>%
  filter(!is.na(hotspot))
@

@@

<<subset_expr_matrix-2>>=
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
nnonlocal <- nrow(nonlocal_annot)
nlocal <- nrow(local_annot2)
@

@@

<<scatter-2>>=
p <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
ggsave(plot = p, filename = file.path("../figures", fnsvg))
ggsave(plot = p, filename = file.path("../figures", fnpdf))
p
@

@@

fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
ggsave(plot = p, filename = file.path("../figures", fnsvg), height = ceiling(nlocal / 2))
ggsave(plot = p, filename = file.path("../figures", fnpdf), height = ceiling(nlocal / 2))

@@

for (i in 1:n_figs){
  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
  foo %>%
    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
    ggplot() +
      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
      broman::karl_theme() +
      theme(legend.position = "none") +
      theme(
        strip.background = element_blank(),
        strip.text.x = element_blank()
    ) +
      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
      ylab("Mediation LOD difference") +
      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
  ggsave(filename = file.path("../figures", fnsvg), height = ceiling(n_panels / 2))
  ggsave(filename = file.path("../figures", fnpdf), height = ceiling(n_panels / 2))
}
@

@@

p <- pm[ord, ord] %>%
  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
  heatmaply(Rowv = FALSE,
            Colv = FALSE,
            cexRow = 0.2,
            cexCol = 0.2,
            symm = TRUE,
            row_side_colors = scols[ord, ],
            col_side_colors = tibble(local = scols1[ord]),
            limits = c(0, 5)
            ) %>%
  hide_guides()
fnsvg <- paste0("Chr", hot_chr, "_heat", ".svg")
fnpdf <- paste0("Chr", hot_chr, "_heat", ".pdf")
plotly::orca(p, file.path("../figures", fnsvg))
plotly::orca(p, file.path("../figures", fnpdf))
p
@

>>>>

<<readRDS-2>>=
pleio_mixed_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio.rds")))
pleio_local_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_local.rds")))
pleio_nonlocal_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_pleio_nonlocal.rds")))
med_list <- readRDS(here("analysis", "data", "derived_data", paste0("Chr", hot_chr, "_mediations.rds")))
@

<<load_attie, eval=!file.exists(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))>>=
download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
              )
load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
@

<<read_attie_files>>=
lod_peaks <- readRDS(here("analysis", "data", "derived_data", "lod_peaks.rds"))
annots <- readRDS(here("analysis", "data", "derived_data", "annots.rds"))
@

<<load_tidyverse>>=
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
@

<<make_local_annot-2>>=
local_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= ult) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
@

<<make_nonlocal_annot-2>>=
nonlocal_annot <- lod_peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  inner_join(y = annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>%
  filter(!is.na(hotspot))
@

@@

<<subset_expr_matrix-2>>=
hot_local <- expr[, colnames(expr) %in% local_annot2$annot.id]
hot_nonlocal <- expr[, colnames(expr) %in% nonlocal_annot$annot.id]
nnonlocal <- nrow(nonlocal_annot)
nlocal <- nrow(local_annot2)
@

@@

<<scatter-2>>=
p <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = local_symbol == keller_mediator, nonlocal = nonlocal_symbol), size = 0.05) + theme(legend.position = "none") + xlab("Pleiotropy LOD") + ylab("Mediation LOD difference")
fnsvg <- paste0("Chr", hot_chr, "_scatter.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter.pdf")
ggsave(plot = p, filename = here("analysis", "figures", fnsvg))
ggsave(plot = p, filename = here("analysis", "figures", fnpdf))
p
@

@@

fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
ggsave(plot = p, filename = here("analysis", "figures", fnsvg), height = ceiling(nlocal / 2))
ggsave(plot = p, filename = here("analysis", "figures", fnpdf), height = ceiling(nlocal / 2))

@@

for (i in 1:n_figs){
  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
  foo %>%
    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
    ggplot() +
      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) +
      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
      broman::karl_theme() +
      theme(legend.position = "none") +
      theme(
        strip.background = element_blank(),
        strip.text.x = element_blank()
    ) +
      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) +
      ylab("Mediation LOD difference") +
      geom_text(size = 2, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
  ggsave(filename = here("analysis", "figures", fnsvg), height = ceiling(n_panels / 2))
  ggsave(filename = here("analysis", "figures", fnpdf), height = ceiling(n_panels / 2))
}
@

@@

p <- pm[ord, ord] %>%
  apply(FUN = trunc2, MARGIN = 2, threshold = 5) %>%
  heatmaply(Rowv = FALSE,
            Colv = FALSE,
            cexRow = 0.2,
            cexCol = 0.2,
            symm = TRUE,
            row_side_colors = scols[ord, ],
            col_side_colors = tibble(local = scols1[ord]),
            limits = c(0, 5)
            ) %>%
  hide_guides()
fnsvg <- paste0("Chr", hot_chr, "_heat", ".svg")
fnpdf <- paste0("Chr", hot_chr, "_heat", ".pdf")
plotly::orca(p, here("analysis", "figures", fnsvg))
plotly::orca(p, here("analysis", "figures", fnpdf))
p
@
